---
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# outputter

The design goal of **outputter** is to provide a simple closure-based syntax for saving/writing output from iterative simulations such that:

- The simulation code can just call a function with data to be written as arguments
- How the results are written/saved (e.g. a data frame object, a text file, a table in a database) can be decided outside of the actual simulation code.

See below for examples.




## Installation

Install development version from GitHub with

```{r install-github, eval=FALSE}
devtools::install_github("mbojan/outputter")
```







## Examples of use

A common problem when writing iterative simulations is implementing several options of saving the results. Ideally one would like to be able to "factor-out" the actual backend, i.e. how the results are saved, outside of the simulation code. This is what **outputter** attempts to achieve with a concept of "output sinks".



### Basic use


### An implementation of a mock-up simulation function

Consider the following function performing a mock-up iterative simulation:

```{r run1}
run <- function(niter=5) {
  a <- 0
  b <- 5
  for( i in seq(1, niter) ) {
    # Here a lot of complex code producing output as two scalars `a` and `b`
    a <- a + i + 1
    b <- b + i - 1
  }
  list(a=a, b=b)
}
```

The function will go through the `for` loop and return the final values of `a` and `b` as a list.

```{r}
run()
```

What we would like instead is to:

- Store the values of `a` and `b` from all simulation steps.
- Have a choice how the values are stored.
- Don't have to modify the original simulation code with any output-processing code too much.

We can update `run()` from above by adding

1. the `out` argument.
2. a short code interpreting `out` argument that creates the output sink called `out()`.
3. a call to `out()` that "spits" the results out from within the loop.
4. a final call to `out()` that returns the results or some object that can be used to retrieve them.

The original simulation code (the body of the `for` loop) remains almost untouched.

```{r run2}
run <- function(niter=5, out=NULL) {
  # Create output sink function `out()` interpreting the `out` argument
  out <- switch(
    data.class(out),
    "function" = out, # `out` can be user-submitted f(...)
    outputter::make_output(out) # create a sink based on provided object
  )
  
  # The actual simulation code
  a <- 0
  b <- 5
  for( i in seq(1, niter) ) {
    # Here a lot of complex code producing output as two scalars `a` and `b`
    a <- a + i + 1
    b <- b - i - 1
    out(iteration=i, a=a, b=b)
  }
  
  out()
}
```

That solves the following usecases:

1. Output as a data frame
    ```{r}
    run()
    ```

2. Append an existing data frame
    ```{r}
    d <- data.frame(a=-1, b=-1)
    run(out=d)
    ```
    Notice a new column `iteration` has been added.

3. Write to a connection
    ```{r}
    # Setup the connection
    outfile <- tempfile()
    con <- file(outfile, open="w")
    
    # Run
    run(out=con)
    
    # Lines of output
    readLines(outfile)
    
    # Cleanup
    close(con)
    unlink(outfile)
    ```

4. Append a table in a database (needs packages `DBI`, `RSQLite` and `dplyr`)
    ```{r}
    # Set-up a table in an in-memory SQLite database
    con <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")
    # An empty table in database
    DBI::dbWriteTable(con, "output", data.frame(iteration=numeric(0), a=numeric(0), b=numeric(0)))
    # A tbl_dbi table "pointing to" the database table
    d <- dplyr::tbl(con, "output")
    
    # Run!
    run(out=d)   # returns a tbl_dbi object
    
    # The results can be collected from `d` too
    dplyr::collect(d)
    
    # Cleanup
    DBI::dbDisconnect(con)
    rm(con, d)
    ```

